---
title: "Quant13B - APM 2024 paper, mirt"
date: "2024-08-15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quantshort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```
  
```{r setup}
library(haven)
library(mirt)
library(tidyverse)
library(patchwork)

library(TestGardener)
source("R/nominal_surpmat.R")
```

## Read and prepare data

```{r}
n    <- 80
data <- read.csv("data/Quant_13B_sample.csv", header = T) 
N    <- nrow(data) # Number of examinees
U <- as.matrix(unname(data))

# read the key
knostring <- scan("data/Quant_13B_key.txt", "o") 
key       <- as.integer(unlist(stringr::str_split(knostring,"")))

# Number of options per item
noption <- rep(4,n)
noption[c(23:28,63:68)] <- 5
```

```{r}
# Set missing/spoiled responses as NA
# apply(U, 2, table)

for (i in 1:n) {
  U[which(U[,i] > noption[i]), i] <- NA
}

#apply(U, 2, table)

colnames(U) <- paste0("Item", 1:n) # `mirt` requires that items must have unique names in data input
```

## Fit the mirt nominal model

```{r}
file_name <- "data/saved results/mirt_mod_13.rds"
if (file.exists(file_name)) {
  mirt_mod_13 <- readRDS(file_name)
} else {
  start <- Sys.time()
  mirt_mod_13 <- mirt(U, model = 1, itemtype = "nominal", SE = FALSE)
  print(paste0("Run time: ", Sys.time() - start))
  
  saveRDS(mirt_mod_13, file=file_name)
}
```

## Extract parameters: 

This extracts slopes and intercepts for all 80 items. For the four-option items the first 4 in each line are the first 4 slopes, the next 4 the corresponding intercepts and the last two are NA since they aren't used in the nominal model.  For the five-item items the 9th value is the slope and the 10th the intercept for option 5.

```{r}
item_params <- coef(mirt_mod_13, IRTpars = T, simplify = T)$items
```

## Set up for TG style plotting

```{r}
n <- 80
noption <- rep(4,n)
noption[c(23:28,63:68)] <- 5
```

```{r}
ScoreList <- vector("list",n) # option scores
for (item in 1:n){
  scorei <- rep(0,noption[item])
  scorei[key[item]] <- 1
  ScoreList[[item]] <- scorei
}
itemLab <- NULL  #  no item labels
optLab  <- NULL  #  no option labels
optList <- list(itemLab, optLab, optScr=ScoreList)
```

## Nominal

```{r}
titlestr  <- "SweSAT-Q_13B_nominal"
Quant_13B_dataList_nom <- list(optList=optList, key=key, titlestr)
saveRDS(Quant_13B_dataList_nom, file="data/saved results/Quant_13B_dataList_nom.rds")
```

### Set up object WfdList_nom: here the objects for each item to be stored in a named list object WfdListi

```{r}
WfdList_nom <- vector("list", n)
# Specify here the range of theta to consider (from -inf to inf is not possible since probs go to 0)
bdry <- c(-2.5,4)
indfine <- seq(bdry[1],bdry[2],len=101)
for (item in 1:n) {
  if (noption[item] == 4) {
    parmat <- matrix(item_params[item,1:8],4,2)
  } else {
    parmat <- matrix(c(item_params[item,1:4],
                       item_params[item,9],
                       item_params[item,5:8],
                       item_params[item,10]),5,2)
  }
  result <- nominal_surpmat(parmat, bdry, nderiv=2) 
  WListi <- list(
    #  Wfd is a matrix of slopes and intercepts
    Wfd = parmat,
    M = noption[item],
    Pbin = NULL,
    Wbin = NULL,
    Pmatfine = result$prbmat,
    Wmatfine = result$surpmat,
    DWmatfine = result$Dthetasurpmat,
    D2Wmatfine = result$D2thetasurpmat,
    PStdErr = NULL,
    WStdErr = NULL
  )  
  WfdList_nom[[item]] <- WListi
}

surpmat <- result$surpmat
```

### compute score index values (thetas) using EAP method

```{r}
theta_nom <- fscores(mirt_mod_13, "EAP")
```

###  Quant_13B_parList_nom

```{r}
logdensbasis <- fda::create.bspline.basis(bdry, 15)
theta.distnList <- theta.distn(theta_nom, logdensbasis)
pdf_fd    <- theta.distnList$pdf_fd
logdensfd <- theta.distnList$logdensfd
cdffine   <- theta.distnList$cdffine
C         <- theta.distnList$C
denscdf   <- theta.distnList$denscdf
indcdf    <- theta.distnList$indcdf
Pvec      <- c(5,25,50,75,95)/100
Qvec_nom  <- pracma::interp1(as.numeric(denscdf), 
                             as.numeric(indcdf), Pvec)

Quant_13B_parList_nom <- list(
  theta_nom   = theta_nom,
  WfdList_nom = WfdList_nom,
  logdensfd   = logdensfd,
  denscdf     = denscdf,
  indcdf      = indcdf,
  Qvec_nom    = Qvec_nom
)

saveRDS(Quant_13B_parList_nom, file="data/saved results/Quant_13B_parList_nom.rds")
```

###  Quant_13B_infoList_nom

```{r}
file_name <- "data/saved results/Quant_13B_infoList_nom.rds"
if (file.exists(file_name)) {
  Quant_13B_infoList_nom <- readRDS(file_name)
} else {
  Quant_13B_infoList_nom <- theta2arclen(theta_nom, Qvec_nom, WfdList_nom, NULL,
                                             plotrng=bdry) 
  
  saveRDS(Quant_13B_infoList_nom, file=file_name)
}
```

```{r}

```







